Fast Hyper-Parameter Selection for Graph Kernels via Subsampling and Multiple Kernel Learning

### 2016-11-24 paper.tex
Added dataset description

% My packages
\usepackage{ctable} % for \specialrule command
%\usepackage{tikz}
%\usepackage{pgfplots}
%\usepackage{filecontents}
%\usepackage{amsfonts}
%\usepackage{booktabs}
%\usepackage{amsthm}
\usepackage{amssymb,amsmath,array}
%\usepackage{color}
%\usepackage{algorithm}
%\usepackage{algorithmic}
%\usepackage{multirow}
%\usepackage{array}
%\usepackage{subfig}
%\usepackage{caption}
%\usepackage{graphicx}
%\usepackage{hyperref}
%\usepackage{algorithmicx}
%\usepackage{float}
%\usepackage{soul}
%\usepackage{amsfonts}
%\usepackage{subfigure}
\usepackage{amsbsy}
%\usepackage{setspace}
%\usepackage{hyperref}
%\usepackage[bookmarks=false]{hyperref}
%\usepackage{babel,blindtext}
\usepackage{color}

@@

\begin{document}
%style file for ESANN manuscripts
\title{Title}

%***********************************************************************
% AUTHORS INFORMATION AREA
%***********************************************************************
%\author{First author$^1$ and Second author$^2$
\author{A$^1$, B$^2$ and C$^3$
%
% Optional short acknowledgment: remove next line if non-needed
%\thanks{This is an optional funding source acknowledgement.}
%
% DO NOT MODIFY THE FOLLOWING '\vspace' ARGUMENT
\vspace{.3cm}\\
%
% Addresses and institutions (remove "1- " in case of a single institution)
1- Department of Mathematics, University of Padova\\
via Trieste 63, Padova, Italy
\vspace{.1cm}\\
2- Computational Statistics and Machine Learning (CSML) \\
Istituto Italiano di Tecnologia, Via Morego 30, Genova, Italy
\vspace{.1cm}\\
3- ... \\
...
}

@@

%\section{Kernel for structures and Multiple Kernel Learning}
\section{Background}
A kernel method is composed by the following two modules.
A learning algorithm, that expresses the solution only via dot products between training examples.
A kernel function ${k}: \mathcal{X} \times \mathcal{X} \rightarrow \mathbb{R}$ that is a symmetric positive semi-definite function that corresponds to a dot product in a Reproducing Kernel Hilbert Space (RKHS), i.e. there exists a $\phi: \mathcal{X} \rightarrow \mathcal{K} \subseteq \mathbb{R}^D$, where $\mathcal{X}$ is the input space and $\mathcal{K}$ is an Hilbert space (commonly referred to as \textit{feature space}), such that $k(X_i,X_j) = \langle \phi(X_i),\phi(X_j) \rangle$ with $X_i,X_j \in \mathcal{X}$.%~\cite{559923}.
\textcolor{red}{TODO definire kernel matrix}
When dealing with structured data, the choice of the kernel to adopt is a key step. Depending on the type of data under consideration, different proposals are available in literature, each one with a different trade-off between predictive performance and computational time (of course, this trade-off depends from the considered task).

@@

The Weisfeiler-Lehman Fast Subtree kernel (WL)~\cite{NIPS2009_0533}  counts the number of identical subtree patterns obtained by subtree-walks up to height $h$ in the two input graphs. The complexity of the kernel is $O(|E|h)$. 
%While being fast to compute, the kernel may lack of expressiveness for some tasks given that the number of non-zero features generated by one graph is at most $|V|h$. 
Note that the subtree-walks extracted by the kernel differ from  subtree structures because a node can appear multiple times in the same subtree-walk.

In~\cite{Costa2010} the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK), an extension of the WL Fast Subtree Kernel, is presented. This kernel computes the exact matches between pairs of subgraphs with controlled size and distance. 
Their kernel, has $O(|V| |V_h| |E_h| \log |E_h|)$ time complexity, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
% The complexity of the kernel is $O(|V| |V_h| |E_h| \log |E_h|)$, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
The authors state that, for small values of the subgraph size and distance, the complexity of the kernel becomes in practice linear.

The Ordered Decomposition DAGs kernel framework (ODD)~\cite{Dasan2012} considers as non-zero features in the RKHS the trees that appear as subtrees of the input graphs.
It exploits the shortest-path (up to length $\eta$) DAG decompositions starting from each node in the graph to generate DAG structures, and then extracts tree features from them.
Each tree-feature is weighted as $f \cdot \lambda^{dim}$, where $f$ is the frequency of the feature, $dim$ its dimension (the number of vertices in the tree) and $\lambda>0$ a weighting parameter.
The time complexity of the main representative of this family, ODD$_{ST}$, is, under mild conditions, $O(|V_G| \log |V_G|)$.%, and the number of generated features for a graph is at most $|V_G|\rho^\eta$.

@@

\section{Experimental Assessment}
Firstly, we performed some experiments to observe how many samples are required to make $\eta$ and $\eta_s$ similar.
The metric used for comparison is the jaccard similarity score. Indeed, in our context two weights vectors are similar is they provide the same top kernels.
These experiments are summarized in \figurename\ \ref{fig:weights} and \ref{fig:jaccard}.

>>>>

% My packages
\usepackage{ctable} % for \specialrule command
%\usepackage{tikz}
%\usepackage{pgfplots}
%\usepackage{filecontents}
%\usepackage{amsfonts}
%\usepackage{booktabs}
%\usepackage{amsthm}
\usepackage{amssymb,amsmath,array}
\usepackage[utf8]{inputenc}
%\usepackage{color}
%\usepackage{algorithm}
%\usepackage{algorithmic}
%\usepackage{multirow}
%\usepackage{array}
%\usepackage{subfig}
%\usepackage{caption}
%\usepackage{graphicx}
%\usepackage{hyperref}
%\usepackage{algorithmicx}
%\usepackage{float}
%\usepackage{soul}
%\usepackage{amsfonts}
%\usepackage{subfigure}
\usepackage{amsbsy}
%\usepackage{setspace}
%\usepackage{hyperref}
%\usepackage[bookmarks=false]{hyperref}
%\usepackage{babel,blindtext}
\usepackage{color}

@@

\begin{document}
%style file for ESANN manuscripts
%primo tentativo
\title{Fast Hyper-Parameter Selection for Graph Kernels via Subsampling and Multiple Kernel Learning}

%***********************************************************************
% AUTHORS INFORMATION AREA
%***********************************************************************
%\author{First author$^1$ and Second author$^2$
\author{Michele Domini$^1$, Nicol\`o Navarin$^2$, Ivano Lauriola$^2$, Fabio Aiolli$^2$ and Fabrizio Costa$^3$
%
% Optional short acknowledgment: remove next line if non-needed
%\thanks{This is an optional funding source acknowledgement.}
%
% DO NOT MODIFY THE FOLLOWING '\vspace' ARGUMENT
\vspace{.3cm}\\
%
% Addresses and institutions (remove "1- " in case of a single institution)
1- Computational Statistics and Machine Learning (CSML) \\
Istituto Italiano di Tecnologia, Via Morego 30, Genova, Italy
\vspace{.1cm}\\
2- Department of Mathematics, University of Padova\\
via Trieste 63, Padova, Italy
\vspace{.1cm}\\
3- ... \\
...
}

@@

%\section{Kernel for structures and Multiple Kernel Learning}
\section{Background}
A kernel method is composed by the following two modules.
A learning algorithm, that expresses the solution only via dot products between training examples.
A kernel function ${k}: \mathcal{X} \times \mathcal{X} \rightarrow \mathbb{R}$ that is a symmetric positive semi-definite function that corresponds to a dot product in a Reproducing Kernel Hilbert Space (RKHS), i.e. there exists a $\phi: \mathcal{X} \rightarrow \mathcal{K} \subseteq \mathbb{R}^D$, where $\mathcal{X}$ is the input space and $\mathcal{K}$ is an Hilbert space (commonly referred to as \textit{feature space}), such that $k(x_i,x_j) = \langle \phi(x_i),\phi(x_j) \rangle$ with $x_i,x_j \in \mathcal{X}$. %~\cite{559923}.
%\textcolor{red}{TODO definire kernel matrix}
A kernel (or \textit{Gram}) matrix for a set of examples $Tr=\{x_1, \ldots, x_l\}$ is an $l \times l$ matrix $G$ with entries $\{G\}_{ij}=k(x_i,x_j)$.
When dealing with structured data, the choice of the kernel to adopt is a key step. Depending on the type of data under consideration, different proposals are available in literature, each one with a different trade-off between predictive performance and computational time (of course, this trade-off depends from the considered task).

@@

The Weisfeiler-Lehman Fast Subtree kernel (WL)~\cite{NIPS2009_0533}  counts the number of identical subtree patterns obtained by subtree-walks up to height $h$ (an hyper-parameter) in the two input graphs. The complexity of the kernel is $O(|E|h)$. 
%While being fast to compute, the kernel may lack of expressiveness for some tasks given that the number of non-zero features generated by one graph is at most $|V|h$. 
Note that the subtree-walks extracted by the kernel differ from  subtree structures because a node can appear multiple times in the same subtree-walk.

In~\cite{Costa2010} the Neighborhood Subgraph Pairwise Distance Kernel (NSPDK), an extension of the WL Fast Subtree Kernel, is presented. This kernel computes the exact matches between pairs of subgraphs with controlled size and distance (hyper-parameters $r$ and $d$, respectively). 
The kernel has $O(|V| |V_h| |E_h| \log |E_h|)$ time complexity, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
% The complexity of the kernel is $O(|V| |V_h| |E_h| \log |E_h|)$, where $|V_h|$ and $|E_h|$ are the number of nodes and the number of edges of the subgraph obtained by a breadth-fist visit of depth $h$. 
The authors state that, for small values of the subgraph size and distance, the complexity of the kernel becomes in practice linear.

The Ordered Decomposition DAGs Subtree kernel (ODD$_{ST}$)~\cite{Dasan2012,DaSanMartino2016} considers as non-zero features in the RKHS the trees that appear as subtrees of the input graphs.
It exploits the shortest-path (up to length $h$, a hyper-parameter) DAG decompositions starting from each node in the graph to generate DAG structures, and then extracts tree features from them.
Each tree-feature is weighted as $f \cdot \lambda^{dim}$, where $f$ is the frequency of the feature, $dim$ its dimension (the number of vertices in the tree) and $\lambda>0$ a weighting hyper-parameter.
The time complexity of the kernel is, under mild conditions, $O(|V_G| \log |V_G|)$.%, and the number of generated features for a graph is at most $|V_G|\rho^\eta$.

@@

\section{Experimental Assessment}
We conducted our experiments on five real-world datasets:
AIDS~\cite{Weislow1989a}, CAS\footnote{http://cheminformatics.org/datasets/bursi/},
CPDB~\cite{Helma2004a}, NCI1~\cite{Wale2006} and GDD~\cite{dobson2003}.
The first four datasets contain chemical particles encoded in graph form, while the latter encodes proteins.
All the nodes are labelled and there are no self loops.
The AIDS dataset contains 1503 chemical compounds, labelled according to
their antiviral activity; CAS and CPDB store 4337 and 684 mutagenic
compounds respectively; NCI1 consists of 4110 chemical compounds screened for activity against 
non-small lung cancer cells; GDD is composed of 1178 X-ray crystal structures of
proteins.

Firstly, we performed some experiments to observe how many samples are required to make $\eta$ and $\eta_s$ similar.
The metric used for comparison is the jaccard similarity score. Indeed, in our context two weights vectors are similar is they provide the same top kernels.
These experiments are summarized in \figurename\ \ref{fig:weights} and \ref{fig:jaccard}.
